### 2017-03-04 A New Hope/Catherine_Analysis.Rmd
Create Catherine_Analysis.Rmd

---
title: "Mines Accident Prediction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mine Accidents
Data sources

```{r Packages}
require(dplyr)
library(RcppRoll)
library(zoo)
```

```{r loading raw data}
setwd("C:/Users/CATHY/OneDrive/Documents/2016-2017 Junior/15 Mines Research/violation-data-analysis/")

if(!exists("accidents")) {
  accidents <- load("./Analysis/data/mshs_accidents.RData")
}
if(!exists("violations")) {
  violations <- load("./Analysis/data/mshs_violations.RData")
}
if(!exists("assessed_violations")) {
  assessed_violations <- load("./Analysis/data/mshs_assssd_violation.RData")
}
```

```{r accident cleaning}
actual.accidents <- accidents %>% filter(is.element(inj_degr_desc, c("DAYS AWAY FROM WORK ONLY", 
                                                                     "DYS AWY FRM WRK & RESTRCTD ACT", 
                                                                     "DAYS RESTRICTED ACTIVITY ONLY", 
                                                                     "FATALITY", 
                                                                     "PERM TOT OR PERM PRTL DISABLTY", 
                                                                     "NO VALUE FOUND")))
na.and.no.value.index <- is.element(actual.accidents$inj_degr_desc, "NO VALUE FOUND") & is.na(actual.accidents$days_lost)
actual.accidents[na.and.no.value.index, "days_restrict"] <- 0
actual.accidents[na.and.no.value.index, "days_lost"] <- 0
actual.accidents[which(actual.accidents$inj_degr_desc == "FATALITY"), "days_lost"] <- 300
actual.accidents$ai_dt_actual_date <- strftime(actual.accidents$ai_dt, "%F")

day.level.accidents <- actual.accidents %>% group_by(mine_id, ai_dt) %>% summarize(day_level_days_lost = sum(days_lost, na.rm = T))
quarter.level.accidents <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr) %>% summarize(quarter_level_days_lost = sum(days_lost, na.rm = T))
```

```{r accident calculation}

longest.period <- 3
actual.start.year <- 2000
start.year <- 2000 - longest.period
mines.with.accidents.post.start.year <- (actual.accidents %>% filter(cal_yr >= actual.start.year))$mine_id %>% unique()
end.year <- 2015
number.years <- end.year - start.year + 1

roll_over <- function(longest.period, actual.start.year, start.year, )
## Challenge: fill in nonpresent values for mines with 0 days lost
## For example, mine could have just not had a day lost, but that won't show up in `quarter.level.accidents`
## 1. First, initialize all blank data frame by figuring out final structure (we know exactly what pattern mine_id, quarter, and year should follow
##    a particular set of repetitions)
## 2. Anti-join will only keep data from the set of keys not found in the actual data
## 3. Then we can just join the two sets of rows!
blank.full.quarter.level.days_lost <- data.frame(mine_id = rep(mines.with.accidents.post.start.year, each = 4 * number.years),
                                                 quarter = rep(1:4, times = length(mines.with.accidents.post.start.year) * number.years),
                                                 year = rep(rep(start.year:end.year, times = rep(4, number.years)),
                                                            times = length(mines.with.accidents.post.start.year)),
                                                 num.days.lost = 0)
full.quarter.level.days_lost <- anti_join(blank.full.quarter.level.days_lost, quarter.level.accidents,
                                          by = c("mine_id" = "mine_id", "quarter" = "cal_qtr", "year" = "cal_yr"))
full.quarter.level.days_lost <- rbind(full.quarter.level.days_lost,
                                      quarter.level.accidents %>% 
                                        select(mine_id, quarter = cal_qtr,
                                               year = cal_yr, num.days.lost = quarter_level_days_lost) %>% 
                                        data.frame() %>% filter(year >= start.year) %>% 
                                        filter(year <= end.year)) %>% arrange(mine_id, year, quarter)

## Challenge: Calculating past period statistics
## lag will put value of index n - 1 at index n
## roll_sum will do a sum from the specified number of previous rows and put NA's at the beginning
## We filter with actual.start.year so there are no NA's. 
##   (The year that data is ready and the year that data needs to start being used are different.)
full.quarter.level.days_lost <- full.quarter.level.days_lost %>% 
                                group_by(mine_id) %>%
                                mutate(last.quarter = lag(num.days.lost),
                                       last.year = roll_sum(lag(num.days.lost), 4, align = "right", fill = NA),
                                       last.three.years = roll_sum(lag(num.days.lost), 12, align = "right", fill = NA)) %>%
                                filter(year >= actual.start.year)

full.quarter.level.days_lost
#664128
```
```{r violation cleaning}
violations <- mshs_violations
quarter.level.violations <- violations %>% group_by(mine_id, cal_qtr, cal_yr) %>% 
                                            summarize(quarter_level_violation = n())

longest.period <- 3
actual.start.year <- 2000
start.year <- 2000 - longest.period
mines.with.violation.post.start.year <- (violations %>% filter(cal_yr >= actual.start.year))$mine_id %>% unique()
end.year <- 2015
number.years <- end.year - start.year + 1

## Challenge: fill in nonpresent values for mines with 0 days lost
## For example, mine could have just not had a day lost, but that won't show up in `quarter.level.accidents`
## 1. First, initialize all blank data frame by figuring out final structure (we know exactly what pattern mine_id, quarter, and year should follow
##    a particular set of repetitions)
## 2. Anti-join will only keep data from the set of keys not found in the actual data
## 3. Then we can just join the two sets of rows!
blank.full.quarter.level.violations <- data.frame(mine_id = rep(mines.with.violation.post.start.year, each = 4 * number.years),
                                                 quarter = rep(1:4, times = length(mines.with.violation.post.start.year) * number.years),
                                                 year = rep(rep(start.year:end.year, times = rep(4, number.years)),
                                                            times = length(mines.with.violation.post.start.year)),
                                                 viol.quantity = 0)
full.quarter.level.violations <- anti_join(blank.full.quarter.level.violations, quarter.level.violations,
                                          by = c("mine_id" = "mine_id", "quarter" = "cal_qtr", "year" = "cal_yr"))
full.quarter.level.violations <- rbind(full.quarter.level.violations,
                                       quarter.level.violations %>% 
                                        select(mine_id, quarter = cal_qtr,
                                               year = cal_yr, viol.quantity = quarter_level_violation) %>% 
                                        data.frame() %>% filter(year >= start.year) %>% 
                                        filter(year <= end.year)) %>% arrange(mine_id, year, quarter)

## Challenge: Calculating past period statistics
## lag will put value of index n - 1 at index n
## roll_sum will do a sum from the specified number of previous rows and put NA's at the beginning
## We filter with actual.start.year so there are no NA's. 
##   (The year that data is ready and the year that data needs to start being used are different.)
full.quarter.level.violations <- full.quarter.level.violations %>% 
  group_by(mine_id) %>%
  mutate(last.quarter = lag(viol.quantity),
         last.year = roll_sum(lag(viol.quantity), 4, align = "right", fill = NA),
         last.three.years = roll_sum(lag(viol.quantity), 12, align = "right", fill = NA)) %>%
  filter(year >= actual.start.year)

full.quarter.level.violations

```
```{r Compiling data}
quarter.days_lost.violation <- merge(full.quarter.level.days_lost, full.quarter.level.violations , by = c("mine_id","year","quarter"),all=TRUE)
quarter.days_lost.violation[is.na(test)] <- 0
quarter.days_lost.violation
```

```{r Saving Data into RData files}
write.csv(full.quarter.level.days_lost, "./A New Hope/quarter_level_days_lost.csv")
write.csv(full.quarter.level.violations, "./A New Hope/quarter_level_violations.csv")
write.csv(quarter.days_lost.violation, "./A New Hope/quarter_days_lost_violation.csv")

save(full.quarter.level.days_lost,file = "./A New Hope/quarter_level_days_lost.RData")
save(full.quarter.level.violations, file = "./A New Hope/quarter_level_violations.RData")
save(quarter.days_lost.violation, file = "./A New Hope/quarter_days_lost_violation.RData")
```

```{r}
test <- quarter.days_lost.violation %>% select(num.days.lost,last.quarter.x,last.year.x,last.three.years.x,last.quarter.y,last.year.y, last.three.years.y)

pairs(test)
```
removing rows containing more than 50% zeros 
```{r}

reduced_both <- lm(num.days.lost~last.quarter.x
                 +last.year.x
                 +last.three.years.x
                 +last.quarter.y
                 +last.year.y
                 +last.three.years.y,
                 data = quarter.days_lost.violation)
summary(reduced_both)
```
```{r}
reduce_days_lost <- lm(num.days.lost~last.quarter.x
                 +last.year.x
                 +last.three.years.x,
                 data = quarter.days_lost.violation)
summary(reduce_days_lost)
```

```{r}
reduce_days_lost <- lm(num.days.lost~factor(last.year.x),
                 data = quarter.days_lost.violation)
summary(reduce_days_lost)

plot(num.days.lost~last.quarter.x,data=quarter.days_lost.violation)
```

```{r}
temp <- quarter.days_lost.violation %>% filter(last.quarter.x != 0, last.quarter.y != 0,
                                                 last.year.x != 0, last.year.y != 0,
                                                 last.three.years.x != 0, last.three.years.y != 0)
plot(num.days.lost~last.quarter.x,data=temp)

reduced_both_remove_0 <- lm(num.days.lost~last.quarter.x
                 +last.year.x
                 +last.three.years.x
                 +last.quarter.y
                 +last.year.y
                 +last.three.years.y,
                 data = temp)
summary(reduced_both_remove_0)

```

```{r}
over_200 <- quarter.days_lost.violation %>% filter(last.quarter.x > 200)
under_200 <- quarter.days_lost.violation %>% filter(last.quarter.x < 200)

#historgram of over 200 days lost from last year

qplot(num.days.lost, data=over_200, geom="histogram")
plot(num.days.lost ~ last.quarter.x,data = over_200)
plot(num.days.lost ~ last.quarter.x,data = under_200)

```

```{r roll_over}
roll_over <- function(data, quarter.level, longest.period, actual.start.year,end.year){
  start.year <- actual.start.year-longest.period

  number.years <-  end.year - start.year + 1

  mines.with.accidents.post.start.year <- (data %>% filter(cal_yr >= actual.start.year))$mine_id %>% unique()

  blank.full.quarter.level.days_lost <- data.frame(mine_id = rep(mines.with.accidents.post.start.year, each = 4 * number.years),
                                                 quarter = rep(1:4, times = length(mines.with.accidents.post.start.year) * number.years),
                                                 year = rep(rep(start.year:end.year, times = rep(4, number.years)),
                                                            times = length(mines.with.accidents.post.start.year)),
                                                 num.days.lost = 0)
  full.quarter.level.days_lost <- anti_join(blank.full.quarter.level.days_lost, quarter.level,
                                            by = c("mine_id" = "mine_id", "quarter" = "cal_qtr", "year" = "cal_yr"))
  full.quarter.level.days_lost <- rbind(full.quarter.level.days_lost,
                                        quarter.level %>% 
                                          select(mine_id, quarter = cal_qtr,
                                                 year = cal_yr, num.days.lost = base) %>% 
                                          data.frame() %>% filter(year >= start.year) %>% 
                                          filter(year <= end.year)) %>% arrange(mine_id, year, quarter)

  ## Challenge: Calculating past period statistics
  ## lag will put value of index n - 1 at index n
  ## roll_sum will do a sum from the specified number of previous rows and put NA's at the beginning
  ## We filter with actual.start.year so there are no NA's. 
  ##   (The year that data is ready and the year that data needs to start being used are different.)
  full.quarter.level.days_lost <- full.quarter.level.days_lost %>% 
                                  group_by(mine_id) %>%
                                  mutate(last.quarter = lag(num.days.lost),
                                         last.year = roll_sum(lag(num.days.lost), 4, align = "right", fill = NA),
                                         last.three.years = roll_sum(lag(num.days.lost), 12, align = "right", fill = NA)) %>%
                                  filter(year >= actual.start.year)

  return(full.quarter.level.days_lost)
}
```

```{r full quarter}

longest.period <- 3
actual.start.year <- 2000
end.year <- 2015

quarter.level.accidents <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
                                            summarize(base = sum(days_lost, na.rm = T))
accident_rollover <- roll_over(actual.accidents,
                               quarter.level.accidents,
                               longest.period,
                               actual.start.year,
                               end.year)
accident_rollover

quarter.level.violations <- violations %>% group_by(mine_id, cal_qtr, cal_yr) %>% 
                                            summarize(base = n())
violation_rollover <- roll_over(violations,
                                quarter.level.violations,
                                longest.period,
                                actual.start.year,
                                end.year)
violation_rollover

all.quarter.level.accidents <- accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
                                            summarize(base = sum(days_lost, na.rm = T) + sum(days_restrict, na.rm = T))
all_accident_rollover <- roll_over(accidents,
                                   all.quarter.level.accidents,
                                   longest.period,
                                   1984,
                                   2016)
all_accident_rollover

all_violation_rollover <- roll_over(accidents,
                                   quarter.level.violations,
                                   longest.period,
                                   1984,
                                   2016)
all_violation_rollover

```

```{r}
quaarter.days_lost.violation <- merge(accident_rollover, violation_rollover, by = c("mine_id","year","quarter"),all=TRUE)
quarter.days_lost.violation[is.na(quarter.days_lost.violation)] <- 0
quarter.days_lost.violation

all_accident_violation_rollover <- merge(all_accident_rollover,all_violation_rollover, by = c("mine_id","year","quarter"),all=TRUE)
all_accident_violation_rollover[is.na(all_accident_violation_rollover)]<-0
all_accident_violation_rollover

#remove start year
for (x in unique(all_accident_rollover$mine_id)){
  temp <- all_accident_violation_rollover %>% filter(mine_id==x)
  i <- 1
  while( sum(temp[i,3:11]) == 0){
    i = i+1
  }
  if(i != 1){
    print(x)
    print(i)
    temp <- temp[-1:i,]
  }
}

  #remove end year
for (mine_id in unique(all_accident_rollover$mine_id)){
  temp <- all_accident_violation_rollover %>% filter(mine_id==x)
  i <- nrow(temp)
  while( sum(temp[i,3:11]) == 0){
    i = i-1
  }
  if(i != nrow(temp)){
    print(x)
    print(i)
    temp <- temp[-i:nrow(temp),]
  }
}

```