### 2017-03-09 San Antonio/top_tens.Rmd
nothing specific

---
title: "Top Tens"
author: "Yu"
date: "3/8/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
rm(list=ls())
require(dplyr)
knitr::opts_knit$set(root.dir = "~/Git/violation-data-analysis")
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r top 10 based on num.death}
load("./San Antonio/output/result.RData")
complete.active.quarters %>% arrange(desc(num.death)) %>% select(mine_id, mine.name, year, quarter, num.death) %>% head(10)
```

```{r top 10 based on num.days.lost}
complete.active.quarters %>% filter(mine_id == 101322) %>% select(mine.name, year, quarter, last.year.lost) %>% head(10)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.