### 2017-03-10 San Antonio/simple_lm.Rmd
simple linear model
y = num.days.lost or num.death

---
title: "Simple Linear Models"
author: "Yu"
date: "3/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
knitr::opts_knit$set(root.dir = "~/Git/violation-data-analysis")
```

## Simple Linear Model

Read these two links first: [How to interpret regression p-values](http://blog.minitab.com/blog/adventures-in-statistics-2/how-to-interpret-regression-analysis-results-p-values-and-coefficients) and
[How to interpret the output of the summary() method for an lm object](http://stats.stackexchange.com/questions/59250/how-to-interpret-the-output-of-the-summary-method-for-an-lm-object-in-r)

- lm on num of days lost or restricted:
```{r echo=FALSE}
load("./San Antonio/output/result.RData")
num.days.lm <- lm(0*num.death+1*num.days.lost+1*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
           +last.quarter.viol+last.year.viol+last.three.years.viol
           +last.quarter.death+last.year.death+last.three.years.death, 
           data=complete.active.quarters %>% filter(active))
summary(num.days.lm)
```
- lm on num of deaths:
```{r echo=FALSE}
summary(lm(num.death~last.three.years.lost+last.quarter.viol+last.year.viol, data=complete.active.quarters %>% filter(active)))
```