### 2017-03-17 San Antonio/markov_chains.Rmd
All markov chain complete except for weighted average

---
title: "markov_chains"
output: html_document
---

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
require(plyr)
require(RcppRoll)
require(markovchain)
require(reshape)
knitr::opts_knit$set(root.dir = "~/Documents/Violation_Project")
setwd("~/Documents/Violation_Project")

if(!exists("accidents") | 
   !exists("full.days_lost.accidents.date") | 
   !exists("roll_over") | 
   !exists("mines")) {
  load("./San Antonio/data/secret_ingredients.RData")
  load("./San Antonio/output/result_select_accidents.RData")
  print("raw RData loaded")
}
```

```{r}

temp <- complete.active.quarters %>% filter(active==TRUE)
set.seed(20)
clustering <- kmeans(temp[, 6:ncol(temp)], 3, nstart=20)

freq_table <- table(clustering$cluster)

```

```{r}
clustering
freq_table

clustering$centers
```

Transformed data to become in a melted format containing quarter-year, mine_id allowing for markov chain analysis
```{r}

temp$cluster <- clustering$cluster
temp$date <- paste(as.character(temp$year), "-" , as.character(temp$quarter))
temp2 <- temp %>% select(mine_id, cluster, date)
melted <- melt(temp2, id=c("mine_id", "date"))
aggregate_cluster_seq <- cast(melted, mine_id~date)
```

Markov Chain stuff
```{r}
markov <- function(x){
  x <- as.integer(x)
  cluster_seq_raw <- x[!is.na(x)]
  if(length(cluster_seq_raw) > 1){
    cluster_sequence <- cluster_seq_raw[2:length(cluster_seq_raw)]
    cur_markov_chain <- markovchainFit(data=cluster_sequence)
    return (list("mine_id" = x[1], "markov_chain" = cur_markov_chain$estimate, "num_rows" = length(cluster_seq_raw)))
  }
  return (list("mine_id" = x[1], "markov_chain" = NA, "num_rows" = 0))
}
```

```{r}
result<-alply(aggregate_cluster_seq, 1, markov)
```

```{r}
num_rows_ind <- sapply(result, "[[", 3)
weights <- num_rows_ind/sum(num_rows_ind)
markov_chains_ind <- sapply(result, "[[", 2)
aggregated_markov_chain <- weighted.mean(markov_chains_ind, weights)
```