### 2017-03-18 San Antonio/slides/presentation_slides.Rmd

---
title: "A Data-Driven Early Warning System for Mining Accident"
shorttitle: "Mining Accident Data Analysis"
author: |
  | Yu Luo, Ashutosh Nanda, Shiva Rajgopal, Vinay Ramesh,
  | Venkat Venkatasubramanian, Zhizun Zhang, and Catherine Zhao
shortauthor: "Luo et al."
institute: "Columbia University"
date: "3/27/2017"
output: 
  beamer_presentation:
    # incremental: true
    template: rmd_template.tex
    fig_height: 4.75
    toc: true
    slide_level: 2
    includes:
      in_header: rmd_header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
require(dplyr)
require(ggplot2)
knitr::opts_knit$set(root.dir = "~/Git/violation-data-analysis")
```

# Introduction: Mine Safety and MSHA Data

<!-- https://www.msha.gov/about/mission -->
## Mine Safety and Health Administration (MSHA)

- Formed in	1977
- Agency of the United States Department of Labor
- Mission
    - Prevent death, illness, and injury from mining
    - Promote safe and healthful workplaces for U.S. miners
    - Develop and enforce safety and health rules
    - Provide technical, educational, and other types of assistance

## Fatality Trend Since 1983

```{r fatality trend}
if(!exists("accidents")) {load("./San Antonio/data/secret_ingredients.RData")}
yearly.death <- accidents %>% 
  mutate(death = ifelse(inj_degr_desc == "FATALITY", 1, 0)) %>%
  group_by(cal_yr) %>%
  summarise(count = sum(death))
qplot(x = cal_yr, y = count, data = yearly.death, geom = c("point", "line"),
      main = "Number of Total Fatalities Due to Accidents",
      xlab = "Year", ylab = "Count")
```

## Upper Big Branch Disaster

# Methods: Data and Model Assumptions

## Primary Datasets

- **Mine accidents**: "msha_accident.csv" (681,386 obs., retrieved 1/26/2017, from <https://enforcedata.dol.gov/views/data_summary.php>)
- **MSHA assessed violations**: "AssessedViolations.csv" (2,169,804 obs., retrieved 12/10/2016, from <https://arlweb.msha.gov/OpenGovernmentData/OGIMSHA.asp>)

## Master Data Table

```{r master data}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
names(complete.active.quarters)
```

## Model

# Results and Discussion

## Number of Days Lost since 2000

```{r num.days.lost trend}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.lost <- complete.active.quarters %>% 
  filter(year >= 2000 & active) %>%
  group_by(year) %>%
  summarise(count = sum(num.days.lost),
            avg = mean(num.days.lost))
qplot(x = year, y = count, data = yearly.lost, geom = c("point", "line"),
      main = "Number of Total Days Lost Due to Accidents",
      xlab = "Year", ylab = "Count")
# qplot(x = year, y = avg, data = yearly.lost, geom = c("point", "line"),
#       main = "Number of Average Days Lost Due to Accidents",
#       xlab = "Year", ylab = "Count")
```

## Number of Violations since 2000

```{r viol.quantity trend}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(year >= 2000 & active) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity),
            avg = mean(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Total Violations",
      xlab = "Year", ylab = "Count")
```

# Conclusion

<!-- USEFUL LINKS:
http://www.sr.bham.ac.uk/~ajrs/R/r-show_data.html
http://rmarkdown.rstudio.com/articles_beamer.html
https://groups.google.com/forum/#!searchin/pandoc-discuss/beamer/pandoc-discuss/XMYG6ewG7a0/Q1LKHaOFBDUJ -->