### 2017-03-20 San Antonio/slides/systemic_risk.tex

\begin{frame}{Systemic Risk}\pause
\begin{columns}[c]
	\column{0.475\linewidth}
		\begin{itemize}[<+->]
		\item Systemic disasters
		\begin{itemize}
			\item SARS (2003)
			\item Northeast Blackout (2003)
			\item Subprime Crisis (2008)
			\item Deepwater Horizon
			%Oil Spill 
			(2010)
		\end{itemize}
		\item Emerging systemic risks
		\begin{itemize}
			\item Climate change
			\item Income/wealth inequality
			\item Cyber-physical security
			\item Technological singularity
		\end{itemize}
		\item Fast-paced and connected
		\item Design complex systems
		\item Analyze systemic risk
		\end{itemize}
	\column{0.525\linewidth}
	   \onslide<3->
	   \includegraphics[width = .5\linewidth]{fig/wide_SARS.png}
	   \onslide<8->
	   \includegraphics[width = .5\linewidth]{fig/wide_polarbear.png}\\%http://i.dailymail.co.uk/i/pix/2014/09/09/1410255561618_wps_15_Spitsbergen_Norway_Polar_.jpg
	   \onslide<4->
	   \includegraphics[width = .5\linewidth]{fig/wide_NEB.png}
	   \onslide<9->
	   \includegraphics[width = .5\linewidth]{fig/wide_99.png}\\%https://c1.staticflickr.com/7/6055/6247188828_305409d56a_b.jpg
	   \onslide<5->
	   \includegraphics[width = .5\linewidth]{fig/wide_FIN.png}
	   \onslide<10->
	   \includegraphics[width = .5\linewidth]{fig/wide_cyber.png}\\%https://c1.staticflickr.com/9/8604/16042227002_1d00e0771d_b.jpg
	   \onslide<6->
	   \includegraphics[width = .5\linewidth]{fig/wide_BP.png}
	   \onslide<11->
	   \includegraphics[width = .5\linewidth]{fig/wide_exmachina.png}%https://c1.staticflickr.com/1/569/20167701293_e3ed4b066f_b.jpg
\end{columns}
\end{frame}