### 2017-03-21 San Antonio/slides/presentation_slides.Rmd
updated slides

---
title: "A Data-Driven Early Warning System for Mining Accident"
shorttitle: "Mining Accident Data Analysis"
author: |
  | Yu Luo, Ashutosh Nanda, Shiva Rajgopal, Vinay Ramesh,
  | Venkat Venkatasubramanian, Zhizun Zhang, and Catherine Zhao
shortauthor: "Luo et al."
institute: |
  | Chemical Engineering, Computer Science, and Business School
  | Columbia University
shortinstitute: "Columbia University"
date: "3/27/2017"
output: 
  beamer_presentation:
    keep_tex: false
    incremental: false
    template: rmd_template.tex
    fig_height: 4.75
    toc: true
    slide_level: 2
    includes:
      in_header: rmd_header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, fig.align = "center")
options(scipen = 1, digits = 2)
require(dplyr)
require(ggplot2)
knitr::opts_knit$set(root.dir = "~/Git/violation-data-analysis")
```

## Complex, Resilient, Intelligent Systems (CRIS Lab)

![CRIS Lab](CrisResearch720.png)

<!-- https://aiche.confex.com/aiche/s17/webprogram/Paper480429.html -->
# Introduction: A Data Approach to Mine Safety

## Systemic Risk

- Systemic disasters
    - SARS (2003)
    - Northeast Blackout (2003)
    - Subprime Crisis (2008)
    - Deepwater Horizon Oil Spill (2010)
- Emerging systemic risks
    - Climate change
    - Income/wealth inequality
    - Cyber-physical security
    - Technological singularity
- Fast-paced and connected
- Design complex systems
- Analyze systemic risk

<!-- 
https://en.wikipedia.org/wiki/Upper_Big_Branch_Mine_disaster
https://arlweb.msha.gov/MEDIA/PRESS/2011/NR111206.asp 
-->
## Upper Big Branch Mine Disaster (2010)

- April 5, 2010, Raleigh County, West Virginia, owned by Massey Energy
- 29 deaths, the worst mining in the United States since 1970
- MSHA cites corporate culture as root cause of Upper Big Branch Mine disaster

## Sago Mine Disaster (2006)

- January 2, 2006, Sago, West Virginia, owned by Anker West Virginia Mining
- 13 miners were trapped for nearly two days; only one survived
- Fatality number was exceeded by the Upper Big Branch Mine disaster
- MSHA reports prior history of safety violations and fatalities

<!-- https://www.msha.gov/about/mission -->
## Mine Safety and Health Administration (MSHA)

- Formed in	1977
- Agency of the United States Department of Labor
- Mission
    - Prevent death, illness, and injury from mining
    - Promote safe and healthful workplaces for U.S. miners
    - Develop and enforce safety and health rules
    - Provide technical, educational, and other types of assistance

## Fatality Trend Since 1983

```{r fatality trend}
if(!exists("accidents")) {load("./San Antonio/data/secret_ingredients.RData")}
yearly.death <- accidents %>% 
  mutate(death = ifelse(inj_degr_desc == "FATALITY", 1, 0)) %>%
  group_by(cal_yr) %>%
  summarise(count = sum(death))
qplot(x = cal_yr, y = count, data = yearly.death, geom = c("point", "line"),
      main = "Number of Total Fatalities Due to Accidents",
      xlab = "Year", ylab = "Count")
```

## Can We Further Improve Mine Safety?

- Process MSHA safety data
- Understand the underlying causal relationships
- Develop early warning systems based on past behaviors
- Credit rating/score analogy
    - Predict default probability within 18 months
    - Accidents: defaults a month or a year prior to application
    - Violations: missed payments, late payments, etc.

# Methods: Data Sources and Model Preliminaries

## Department of Labor Enforcement Data

- Link: <https://enforcedata.dol.gov/views/data_catalogs.php>
- Updated daily or weekly
- Enforecement Data Catalog
    - MSHA Enforcement Data
    - OSHA Enforcement Data
    - Wage and Hour Compliance Action Data
    - OFCCP Compliance Evaluation and Complaint Investigation Data
    - EBSA Enforcement Data

## MSHA Data: Sources

- Mine accidents table: "msha_accident.csv" 
    - 681,386 rows
    - Retrieved 1/26/2017, from <https://enforcedata.dol.gov/views/data_summary.php>
- MSHA assessed violations table: "AssessedViolations.csv" 
    - 2,169,804 rows
    - Retrieved 12/10/2016, from <https://arlweb.msha.gov/OpenGovernmentData/OGIMSHA.asp>

## MSHA Data: Advantages

- Each mine has a unique mine ID, e.g., Upper Big Branch (4608436)
- Rich details, e.g., selected attributes from the accidents table:

```{r msha accidents}
if(!exists("accidents")) {load("./San Antonio/data/secret_ingredients.RData")}
names(accidents[c(1, 14, 5:13, 19:20, 37:59)])
```

## MSHA Data: Challenges

- Missing data
- Inactive mines
- Most data are not numeric
- Relatively few severe accidents ($\sim0.5\%$)

## Model Preliminaries

- Create a master data table with selected attributes
- Plot violation trends for Upper Big Branch Mine and Sago Mine
- Design a predictive model

## Master Data Table

- Combine and summarize accidents/violations based on mines
- 664,128 rows, 10,377 unique mines
- From 2000 to 2015 in quarters
- Each row represents data for a unique combination of mine, year, and quarter
    - e.g., Upper Big Branch Mine in the second quarter of 2010
- Each row contains both current and past information

## Master Data Table

```{r master data}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
names(complete.active.quarters)
```

## Top 10 Fatal Accidents Since 2005

```{r top 10 fatality}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
top.deaths <- complete.active.quarters %>% filter(active & year >= 2005) %>% arrange(desc(num.death), mine.name) %>% select(mine.name, mine_id, year, quarter, num.death)
head(top.deaths, 10)
```

## Violation Trend: Upper Big Branch

```{r upper big branch}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(mine_id == 4608436 & year < 2010) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Yearly Violations",
      xlab = "Year", ylab = "Count")
```

## Violation Trend: Sago Mine

```{r sago}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(mine_id == 4608791 & year < 2006) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Yearly Violations",
      xlab = "Year", ylab = "Count")
```

<!--
## Violation Trend: Crandall Canyon

```{r crandall}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(mine_id == 4201715 & year < 2007) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Yearly Violations",
      xlab = "Year", ylab = "Count")
```

## Violation Trend: Darby Mine No. 1

```{r darby}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(mine_id == 1518185 & year < 2006) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Yearly Violations",
      xlab = "Year", ylab = "Count")
```

## Violation Trend: Gibson

```{r gibson}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(mine_id == 1202215 & year < 2007) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Yearly Violations",
      xlab = "Year", ylab = "Count")
```
-->

## Predictive Model

- Rising violation trends before disasters
- A disaster classifier based on historical data?
- Define a **severe** accident as one with death or permenant disability
- Unbalanced data

```{r severe active}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
severe.summary <- complete.active.quarters %>%
  filter(active) %>%
  mutate(severe = ifelse(
    num.death + num.dis > 0, TRUE, FALSE
  )) %>%
  group_by(severe) %>%
  summarise(n = n()) %>%
  mutate(perc = n / sum(n) * 100)
print(severe.summary)
```

<!-- http://www.stata.com/manuals13/rclogit.pdf -->
## Fixed-Mine Effects

- Biostatisticians and epidemiologists call it "conditional logistic regression" (`survival::clogit`)
- Suitable for panel data (e.g., our master data table)
- Model includes mine-specific but time-invariant variables (e.g., same slope but different intercepts for different mines)
- Logistic regression (for every mine)

\[\Pr(Y=1|\mathbf{X})=F(\mathbf{X}) = \frac{1}{1+e^{-(\alpha+\boldsymbol\beta\mathbf{X})}}\]

- Logistic regression with fixed effects (for the $i$-th mine)

\[\Pr(Y=1|\mathbf{X},i)=F(\mathbf{x},i) = \frac{1}{1+e^{-(\alpha_i+\boldsymbol\beta\mathbf{x})}}\]

# Results and Discussion

## Logistic Regression Without Fixed Effects

- Train and test on all data

```{r glm}
if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
print(glm.performance$table)
print(glm.performance$byClass)
```

## Logistic Regression Without Fixed Effects

- True positives

```{r glm rank death}
if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
head(glm.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10) %>% select(
  mine.name, year, quarter, severe, pred
) %>% mutate (mine.name = strtrim(mine.name, 40))
```

## Logistic Regression Without Fixed Effects

- Top predicted positives (possibly false positives)

```{r glm rank prob}
if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
head(glm.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 10) %>% select(
  mine.name, year, quarter, severe, pred
) %>% mutate (mine.name = strtrim(mine.name, 40))
```

## Logistic Regression with Fixed Effects

- Randomly select half of the data to train and the other half to test

```{r fe.div}
if (!exists("fe.div.result")){load("./San Antonio/output/results_fixed_effects.RData")}
print(fe.div.performance$table)
print(fe.div.performance$byClass)
```

## Logistic Regression with Fixed Effects

- Top true positives (deadliest accidents)

```{r fe.div rank death}
if (!exists("fe.div.result")){load("./San Antonio/output/results_fixed_effects.RData")}
head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10) %>% select(
  mine.name, year, quarter, severe, pred
) %>% mutate (mine.name = strtrim(mine.name, 40))
```

## Logistic Regression with Fixed Effects

- Top predicted positives (possibly false positives)

```{r fe.div rank prob}
if (!exists("fe.div.result")){load("./San Antonio/output/results_fixed_effects.RData")}
head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(prob)), 10) %>% select(
  mine.name, year, quarter, severe, pred
) %>% mutate (mine.name = strtrim(mine.name, 40))
```

<!--
## New Era Mine

- The American Coal Company New Era Mine occupies 6 slots on the top 10 list of most days lost due to accidents

```{r top 10 days lost}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
top.days.lost <- complete.active.quarters %>% filter(active) %>% arrange(desc(num.days.lost), mine.name) %>% select(mine.name, year, quarter, num.days.lost)
head(top.days.lost, 10)
```
-->

## New Era Mine

<!--- Violation trend prior to the first quarter in 2006 where model probability is 0.99-->

```{r new era viol}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(mine_id == 1102752) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Yearly Violations",
      xlab = "Year", ylab = "Count")
```

## New Labels Including Days Lost

- Previously defined severe accidents and days lost $>$ 300

```{r fe.div2}
if (!exists("fe.div2.result")){load("./San Antonio/output/results_fixed_effects.RData")}
print(fe.div2.performance$table)
print(fe.div2.performance$byClass)
```

## New Labels Including Days Lost

- Top true positives (deadliest accidents)

```{r fe.div2 rank death}
if (!exists("fe.div2.result")){load("./San Antonio/output/results_fixed_effects.RData")}
head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10) %>% select(
  mine.name, year, quarter, severe, pred
) %>% mutate (mine.name = strtrim(mine.name, 40))
```

## New Labels Including Days Lost

- Top predicted positives (possibly false positives)

```{r fe.div2 rank prob}
if (!exists("fe.div2.result")){load("./San Antonio/output/results_fixed_effects.RData")}
head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(prob)), 10) %>% select(
  mine.name, year, quarter, severe, pred
) %>% mutate (mine.name = strtrim(mine.name, 40))
```

# Conclusion

## Conclusion

<!-- - Causal relationship between accidents and historical data -->
- Summary
    - Two deadliest mine accidents in the last decade: Upper Big Branch & Sago
    - MSHA

## Appendix: Simple Linear Model

- Adjusted $R^2=0.36$

```{r simple lm of days lost}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
num.days.lm <- lm(num.days.lost~last.quarter.lost+last.year.lost+last.three.years.lost
           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
           +last.quarter.viol+last.year.viol+last.three.years.viol
           +last.quarter.death+last.year.death+last.three.years.death, 
           data=complete.active.quarters %>% filter(active))
summary(num.days.lm)$coefficients
```

<!-- USEFUL LINKS:
http://www.sr.bham.ac.uk/~ajrs/R/r-show_data.html
http://rmarkdown.rstudio.com/articles_beamer.html
https://groups.google.com/forum/#!searchin/pandoc-discuss/beamer/pandoc-discuss/XMYG6ewG7a0/Q1LKHaOFBDUJ -->