### 2017-03-29 San Antonio/slides/presentation_slides.Rmd

---
title: "A Data-Driven Early Warning System for Mining Accidents"
shorttitle: "Mining Accident Data Analysis"
author: |
  | Yu Luo, Ashutosh Nanda, Shivaram Rajgopal, Vinay Ramesh,
  | Zhizun Zhang, Catherine Zhao, and Venkat Venkatasubramanian
shortauthor: "Luo et al."
institute: |
  | Chemical Engineering, Computer Science, and Business School
  | Columbia University
shortinstitute: "Columbia University"
date: "3/27/2017"
output: 
  beamer_presentation:
    keep_tex: true
    incremental: false
    template: rmd_template.tex
    fig_height: 4.75
    # fig_width: 4
    toc: true
    slide_level: 2
    includes:
      in_header: rmd_header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, fig.align = "center")
options(scipen = 1, digits = 2)
require(dplyr)
require(ggplot2)
require(markovchain)
knitr::opts_knit$set(root.dir = "~/Git/violation-data-analysis")
```

## Complex, Resilient, Intelligent Systems (CRIS Lab)

![CRIS Lab](CrisResearch720.png)

<!-- https://aiche.confex.com/aiche/s17/webprogram/Paper480429.html -->
# Mine Safety: A Data-Driven Approach

## Systemic Risk

- Systemic disasters
    - SARS (2003)
    - Northeast Blackout (2003)
    - Subprime Crisis (2008)
    - Deepwater Horizon Oil Spill (2010)
- Emerging systemic risks
    - Climate change
    - Income/wealth inequality
    - Cyber-physical security
    - Technological singularity
- Fast-paced and connected
- Prevent systemic disasters
- Analyze systemic risk: go beyond one-off accidents

<!-- https://en.wikipedia.org/wiki/Upper_Big_Branch_Mine_disaster -->
<!-- https://arlweb.msha.gov/MEDIA/PRESS/2011/NR111206.asp  -->
## Upper Big Branch Mine Disaster (2010)

- April 5, 2010, Raleigh County, West Virginia, owned by Massey Energy
- 29 deaths, the worst mining in the United States since 1970
- MSHA cites corporate culture as root cause of Upper Big Branch Mine disaster

## Sago Mine Disaster (2006)

- January 2, 2006, Sago, West Virginia, owned by Anker West Virginia Mining
- 13 miners were trapped for nearly two days; only one survived
- Fatality number was exceeded by the Upper Big Branch Mine disaster
- MSHA reports prior history of safety violations and fatalities

<!-- https://www.msha.gov/about/mission -->
## Mine Safety and Health Administration (MSHA)

- Formed in	1977, agency of the U.S. Department of Labor
- Mission
    - Prevent death, illness, and injury from mining
    - Promote safe and healthful workplaces for U.S. miners
    - Develop and enforce safety and health rules
    - Provide technical, educational, and other types of assistance
- A constantly improving industry in terms of safety

## Fatality Trend Since 1983

```{r fatality trend}
if(!exists("accidents")) {load("./San Antonio/data/secret_ingredients.RData")}
yearly.death <- accidents %>% 
  mutate(death = ifelse(inj_degr_desc == "FATALITY", 1, 0)) %>%
  group_by(cal_yr) %>%
  summarise(count = sum(death))
qplot(x = cal_yr, y = count, data = yearly.death, geom = c("point", "line"),
      main = "Number of Total Fatalities Due to Accidents",
      xlab = "Year", ylab = "Count")
```

## Can We Further Improve Mine Safety?

- Process MSHA safety data
- Understand the underlying causal relationships
- Develop early warning systems based on past behaviors
- Credit rating/score analogy
    - Predict default probability within 18 months
    - Accidents: defaults a month or a year prior to application
    - Violations: missed payments, late payments, etc.
- Can we develop a "credit score" for mine safety?

# Methods: Data Sources and Model Preliminaries

## Department of Labor Enforcement Data

- Link: <https://enforcedata.dol.gov/views/data_catalogs.php>
- Updated daily or weekly
- Publicly available 
    - Department of Labor: MSHA, OSHA, etc.
    - Other departments: EPA, FDA, DOJ, etc.

    <!-- - MSHA Enforcement Data -->
    <!-- - OSHA Enforcement Data -->
    <!-- - Wage and Hour Compliance Action Data -->
    <!-- - OFCCP Compliance Evaluation and Complaint Investigation Data -->
    <!-- - EBSA Enforcement Data -->
    <!-- - Department of Justice -->
    <!-- - EPA -->
    <!-- - FDA -->

## MSHA Data: Sources

- Mine accidents table: "msha_accident.csv" 
    - 681,386 rows
    - Retrieved 1/26/2017, from <https://enforcedata.dol.gov/views/data_summary.php>
- MSHA assessed violations table: "AssessedViolations.csv" 
    - 2,169,804 rows
    - Retrieved 12/10/2016, from <https://arlweb.msha.gov/OpenGovernmentData/OGIMSHA.asp>

## MSHA Data: Advantages

- Each mine has a unique mine ID, e.g., Upper Big Branch (4608436)
- Rich details: e.g., classification, description, and severity
- Selected attributes from the accidents table (omitting 42 attributes)

```{r msha accidents}
if(!exists("accidents")) {load("./San Antonio/data/secret_ingredients.RData")}
names(accidents[c(1, 14, 19:20, 5:13, 37, 39, 49:50)])
```

## MSHA Data: Challenges

- Missing data, human errors
- No information about inactive/nonoperating mines
- Most data are not numeric
- Lots of zeros, few severe accidents ($\sim0.5\%$)

<!-- ## Model Preliminaries -->

<!-- - Create a master data table with selected attributes -->
<!-- - Plot violation trends for Upper Big Branch Mine and Sago Mine -->
<!-- - Conduct statistical analyses -->

## Consolidated Data

- Group and summarize accidents/violations by mines
- 664,128 rows, 10,377 unique mines
- From 2000 to 2015
- Each row represents data for a unique combination of mine, year, and quarter
    - e.g., Upper Big Branch Mine in the second quarter of 2010
- Each row contains both current and past information
    - i.e., current quarter, past quarter, past year, and past three years

## Consolidated Data

<!-- - Number of days lost (`num.days.lost`), numer of days restricted (`num.days.restrict`), number of deaths (`num.death`), number of permenant disabilities (`num.dis`), and number of violations (`viol.quantity`)  -->

- All 25 attributes of the consolidated data

```{r master data}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
names(complete.active.quarters)
```

## Top 10 Fatal Accidents Since 2005

- Query the consolidated data on the deadliest accidents

```{r top 10 fatality}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
top.deaths <- complete.active.quarters %>% filter(active & year >= 2005) %>% arrange(desc(num.death), mine.name) %>% select(mine.name, mine_id, year, quarter, num.death)
head(top.deaths, 10)
```

- Plot violation trends prior to disasters

## Violation Trend: Upper Big Branch

```{r upper big branch}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(mine_id == 4608436 & year < 2010) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Yearly Violations",
      xlab = "Year", ylab = "Count")
```

## Violation Trend: Sago Mine

```{r sago}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(mine_id == 4608791 & year < 2006) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Yearly Violations",
      xlab = "Year", ylab = "Count")
```

<!--
## Violation Trend: Crandall Canyon

```{r crandall}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(mine_id == 4201715 & year < 2007) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Yearly Violations",
      xlab = "Year", ylab = "Count")
```

## Violation Trend: Darby Mine No. 1

```{r darby}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(mine_id == 1518185 & year < 2006) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Yearly Violations",
      xlab = "Year", ylab = "Count")
```

## Violation Trend: Gibson

```{r gibson}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(mine_id == 1202215 & year < 2007) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Yearly Violations",
      xlab = "Year", ylab = "Count")
```
-->

## Predictive Model

- Rising violation trends before disasters
- Develop a disaster classifier based on historical data
- Define a **severe** accident as one with death or permenant disability
<!-- - Unbalanced data -->

```{r severe active}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
severe.summary <- complete.active.quarters %>%
  filter(active) %>%
  mutate(severe = ifelse(
    num.death + num.dis > 0, TRUE, FALSE
  )) %>%
  group_by(severe) %>%
  summarise(n = n()) %>%
  mutate(perc = n / sum(n) * 100)
print(severe.summary)
```

<!-- http://www.stata.com/manuals13/rclogit.pdf -->
## Fixed-Mine Effects

- Biostatisticians and epidemiologists call it "conditional logistic regression" (`survival::clogit`)
- Suitable for **panel data** (e.g., longitudinal data, our consolidated data)
- Model includes mine-specific biases
- Logistic function (for every mine)

\[\Pr(Y=1|\mathbf{X}) = \frac{1}{1+e^{-(\alpha+\boldsymbol\beta\mathbf{X})}}\]

- Logistic function with fixed effects (for the $i$-th mine)

\[\Pr(Y=1|\mathbf{X},i) = \frac{1}{1+e^{-(\alpha_i+\boldsymbol\beta\mathbf{x})}}\]

# Results and Discussion

## Logistic Regression Without Fixed Effects

- In-sample model

```{r glm}
if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
print(glm.performance$table)
print(c(glm.performance$overall["Accuracy"], glm.performance$byClass[c("Sensitivity", "Specificity", "Precision", "F1")]))
```

- Accuracy = (TP + TN)/(P + N)
- Sensitivity/recall = TP/P
- Specificity = TN/N
- Precision = TP/(TP + FP)
- F1: harmonic mean of sensitivity and precision

## Logistic Regression Without Fixed Effects

- Fails to predict top 10 deadliest disasters

```{r glm rank death}
if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
head(glm.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10) %>% select(
  mine.name, year, quarter, severe, pred
) %>% mutate (mine.name = strtrim(mine.name, 40))
```

## Logistic Regression Without Fixed Effects

- List of false positive predictions based on predicted probability
<!-- - Top predicted positives (possibly false positives) -->

```{r glm rank prob}
if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
head(glm.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 10) %>% select(
  mine.name, year, quarter, severe, pred
) %>% mutate (mine.name = strtrim(mine.name, 40))
```

## Logistic Regression with Fixed Effects

- Out-of-sample model (randomly select half of the data to train and the other half to test)

```{r fe.div}
if (!exists("fe.div.result")){load("./San Antonio/output/results_fixed_effects.RData")}
print(fe.div.performance$table)
print(c(fe.div.performance$overall["Accuracy"], fe.div.performance$byClass[c("Sensitivity", "Specificity", "Precision", "F1")]))
```

## Logistic Regression with Fixed Effects

- Successfully predicts top 10 deadliest disasters

```{r fe.div rank death}
if (!exists("fe.div.result")){load("./San Antonio/output/results_fixed_effects.RData")}
head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10) %>% select(
  mine.name, year, quarter, severe, pred
) %>% mutate (mine.name = strtrim(mine.name, 40))
```

## Logistic Regression with Fixed Effects

- Still has a lot of false positive predictions
- List of false positive predictions based on predicted probability

```{r fe.div rank prob}
if (!exists("fe.div.result")){load("./San Antonio/output/results_fixed_effects.RData")}
head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(prob)), 10) %>% select(
  mine.name, year, quarter, severe, pred
) %>% mutate (mine.name = strtrim(mine.name, 40))
```

- What happened in the New Era Mine?

<!--
## New Era Mine

- The American Coal Company New Era Mine occupies 6 slots on the top 10 list of most days lost due to accidents

```{r top 10 days lost}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
top.days.lost <- complete.active.quarters %>% filter(active) %>% arrange(desc(num.days.lost), mine.name) %>% select(mine.name, year, quarter, num.days.lost)
head(top.days.lost, 10)
```
-->

## New Era Mine

- Among the worst mines by number of days lost due to accidents

```{r new-era-num-days-lost}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
head(
  complete.active.quarters %>% filter(active) %>% arrange(desc(num.days.lost)) %>% select(
    mine.name, year, quarter, num.days.lost
  ) %>% rename(days.lost = num.days.lost)
, 10)
```

- Rising violation trend from 2000 to 2005

## New Era Mine

<!--- Violation trend prior to the first quarter in 2006 where model probability is 0.99-->
<!-- 即便没有灾难，也有趋势，警示作用，后来可能改善了 -->

```{r new era viol}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(mine_id == 1102752) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Yearly Violations",
      xlab = "Year", ylab = "Count")
```

## New Labels Including Days Lost

- Updated severe accident label
    - Previously defined criteria plus days lost $>$ 300
- Redo out-of-sample model

```{r fe.div2}
if (!exists("fe.div2.result")){load("./San Antonio/output/results_fixed_effects.RData")}
print(fe.div2.performance$table)
print(c(fe.div2.performance$overall["Accuracy"], fe.div2.performance$byClass[c("Sensitivity", "Specificity", "Precision", "F1")]))
```

- Worse true positive rate, improved F1 score

## New Labels Including Days Lost

- Successfully predicts 9 out of top 10 deadliest accidents

```{r fe.div2 rank death}
if (!exists("fe.div2.result")){load("./San Antonio/output/results_fixed_effects.RData")}
head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10) %>% select(
  mine.name, year, quarter, severe, pred
) %>% mutate (mine.name = strtrim(mine.name, 40))
```

## New Labels Including Days Lost

- Accidents of the New Era mine are now true positives

```{r fe.div2 rank prob}
if (!exists("fe.div2.result")){load("./San Antonio/output/results_fixed_effects.RData")}
head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(prob)), 10) %>% select(
  mine.name, year, quarter, severe, pred
) %>% mutate (mine.name = strtrim(mine.name, 40))
```

# Conclusion

## Conclusion

<!-- - Causal relationship between accidents and historical data -->
- Summary
    - Two deadliest mine accidents in the last decade: Upper Big Branch & Sago
    - Rich MSHA data that need clean-up
    - Supervised predictive model
- Application
    - "Credit score" for mine safety
    - Regulators, mines, stakeholders
- Future
    - Improve model performance
    - Unsupervised clustering, neural nets, etc.
    - Expand data: OSHA, EPA, etc.

## Appendix: Simple Linear Model

- Adjusted $R^2=0.36$

```{r simple lm of days lost}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
num.days.lm <- lm(num.days.lost~last.quarter.lost+last.year.lost+last.three.years.lost
           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
           +last.quarter.viol+last.year.viol+last.three.years.viol
           +last.quarter.death+last.year.death+last.three.years.death, 
           data=complete.active.quarters %>% filter(active))
summary(num.days.lm)$coefficients
```

## Appendix: Unsupervised Clustering

- Apply $k$-means clustering to consolidated data on all 20 features
- 3 clusters: `low`-risk, `mid`-risk, and `high`-risk
- Selected cluster centers (omitting 17 features)

```{r k-means}
if(!exists("clustering")){load("./San Antonio/output/results_markov.RData")}
sorted.cluster.centers <- clustering$centers[cluster.to.risk.label,c("num.days.lost", "num.days.restrict", "num.death")]#, "num.dis", "viol.quantity")]
rownames(sorted.cluster.centers) <- risk.labels
print(sorted.cluster.centers)
```

- Cluster sizes

```{r k-means sizes}
if(!exists("clustering")){load("./San Antonio/output/results_markov.RData")}
sorted.cluster.size <- matrix(data = clustering$size[cluster.to.risk.label], nrow = 1, ncol = 3)
colnames(sorted.cluster.size) <- risk.labels
rownames(sorted.cluster.size) <- "size"
print(sorted.cluster.size)
```

<!-- https://cran.r-project.org/web/packages/markovchain/vignettes/an_introduction_to_markovchain_package.pdf -->
## Appendix: Markov Chain

- Overall transition matrix

```{r markov-transition}
if(!exists("markov.mine")){load("./San Antonio/output/results_markov.RData")}
print(round(markov.mine[], 3))
```

- Steady-state distribution

```{r markov-steady-state}
if(!exists("markov.mine")){load("./San Antonio/output/results_markov.RData")}
print(round(steadyStates(markov.mine), 3))
```

<!-- USEFUL LINKS:
http://www.sr.bham.ac.uk/~ajrs/R/r-show_data.html
http://rmarkdown.rstudio.com/articles_beamer.html
https://groups.google.com/forum/#!searchin/pandoc-discuss/beamer/pandoc-discuss/XMYG6ewG7a0/Q1LKHaOFBDUJ -->