### 2017-08-07 manuscript/journal_main.tex

\documentclass[journal]{IEEEtran}
\usepackage[]{l16cn}
%\renewcommand\hl[1]{#1} % activate to remove highlights
\makenoidxglossaries
\loadglsentries{definitions}

\begin{document}
\title{\huge A Data-Driven Early Warning System for\\Mining Accidents}
\author{
  Yu Luo, Ashutosh Nanda, Shivaram Rajgopal, Vinay Ramesh, Zhizun Zhang, Catherine Zhao, and 
  Venkat Venkatasubramanian$^{\ast}$%
  \thanks{\bf Last updated on~\today{}}
  \thanks{Yu Luo, 
    Columbia University, Department of Chemical Engineering, yl2750@columbia.edu}
  \thanks{Ashutosh Nanda, 
    Columbia University, Department of Computer Science, an2655@columbia.edu}
  \thanks{Shivaram Rajgopal, 
    Columbia Business School, sr3269@columbia.edu}    
  \thanks{Vinay Ramesh, 
    Columbia University, Department of Computer Science, vrr2112@columbia.edu}    
  \thanks{Zhizun Zhang, 
    Columbia University, Department of Chemical Engineering, zz2216@columbia.edu}
  \thanks{Catherine Zhao, 
    Barnard College, Department of Computer Science, caz2114@barnard.edu}    
  \thanks{Venkat Venkatasubramanian$^{\ast}$
    (to whom correspondence should be addressed),
    Columbia University, Department of Chemical Engineering, venkat@columbia.edu}
}

\maketitle

\begin{abstract}
	\input{abstract}
\end{abstract}

\IEEEpeerreviewmaketitle

\input{text}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}