### 2018-05-05 Houston/src/manuscript_preparation.Rmd
Create new branch for the manuscript

---
title: "Manuscript Preparation"
author: "Albert"
date: "4/13/2018"
output: 
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
options(scipen = 1, digits = 2)
require(dplyr)
require(ggplot2)
require(RcppRoll)
require(markovchain)
require(reshape)
require(survival)
require(caret)
require(e1071)
source("./Houston/src/prepare_violation.R")
source("./Houston/src/roll_over.R")

knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN
```

## Data Preparation

### Load data

Loading raw data.

```{r load data, include=FALSE}
if(!exists("Accidents") | 
   !exists("AssessedViolations") | 
   !exists("Mines")) {
  load("./Houston/data/Accidents.RData")
  load("./Houston/data/AssessedViolations.RData")
  load("./Houston/data/Mines.RData")
  print("raw RData loaded")
}
```

Let's examine the loaded raw datasets.

```{r data preview, include=FALSE}
# Attributes in Accidents Table
head(Accidents)

# Attributes in Violations Table
head(AssessedViolations)

# Attributes in Mines Table
head(Mines)
```

The `result.RData` from San Antonio folder is the same as the `Consolidated.RData` in Houston folder.

```{r load result, include=FALSE}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
print("load previous consolidated data")

colnames(complete.active.quarters)
```

### Data Visualization

If we define severe as death plus permenant disability, then severe composition is shown as following.

```{r severe1}
# options(digits = 1)
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
severe.summary <- complete.active.quarters %>%
  filter(active) %>%
  mutate(severe = ifelse(
    num.death + num.dis > 0, TRUE, FALSE
  )) %>%
  group_by(severe) %>%
  summarise(n = n()) %>%
  mutate(perc = n / sum(n) * 100)
print(severe.summary)
```

If we redefine severe as death or permenant diability greater than 0, or more than 300 days lost or days retrict, then the severe composition is following.
```{r severe2}
# options(digits = 1)
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
severe.summary <- complete.active.quarters %>%
  filter(active) %>%
  mutate(severe = ifelse(
    (num.death + num.dis > 0) | (num.days.lost + num.days.restrict > 300), TRUE, FALSE
  )) %>%
  group_by(severe) %>%
  summarise(n = n()) %>%
  mutate(perc = n / sum(n) * 100)
print(severe.summary)
```

## Model Development

### Linear Regression

Read these two links first: [How to interpret regression p-values](http://blog.minitab.com/blog/adventures-in-statistics-2/how-to-interpret-regression-analysis-results-p-values-and-coefficients) and
[How to interpret the output of the summary() method for an lm object](http://stats.stackexchange.com/questions/59250/how-to-interpret-the-output-of-the-summary-method-for-an-lm-object-in-r)

We first examine the linearity between "number of days lost" and "last three years' number of violations".

```{r check linearity, include=FALSE}
ggplot(data=complete.active.quarters, mapping = aes(x=last.three.years.viol,y=num.days.lost))+geom_point()
```
It turns out that these variables do not show linear correlation, which violates the linear regression assumption.

Linear regression on num of days lost or restricted:

```{r, include=FALSE}
if(!exists("complete_active_quarters")){load("./Houston/output/Consolidated.RData")}
colnames(complete_active_quarters)

num.days.lm <- lm(0*num.death+1*num.days.lost+1*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
           +last.quarter.viol+last.year.viol+last.three.years.viol
           +last.quarter.death+last.year.death+last.three.years.death, 
           data=complete.active.quarters %>% filter(active))
summary(num.days.lm)
```
Linear regression on num of deaths:

```{r echo=FALSE}
summary(lm(num.death~last.three.years.lost+last.quarter.viol+last.year.viol, data=complete.active.quarters %>% filter(active)))
```

### Logistic Regression

```{r logistic reg, include=FALSE}
# label data ----
labeled.data.death <- complete.active.quarters %>% filter(active) %>% mutate(
  severe = ifelse(
    num.death + num.dis > 0, TRUE, FALSE
  )
)
labeled.data.lost <- complete.active.quarters %>% filter(active) %>% mutate(
  severe = ifelse(
    num.death + num.dis > 0 | num.days.lost > 300, TRUE, FALSE
  )
)

# logistic regression without fixed effects ----
# regular logistic regression model
glm.model <- glm(severe ~ 
                   last.quarter.lost+last.year.lost+last.three.years.lost
                 +last.quarter.restrict+last.year.restrict+last.three.years.restrict
                 +last.quarter.viol+last.year.viol+last.three.years.viol
                 +last.quarter.death+last.year.death+last.three.years.death,
                 data = labeled.data.death, family = binomial(link='logit'))
summary(glm.model)

glm.prediction <- predict(object = glm.model, type = "response")
glm.result <- labeled.data.death %>% select(mine_id, mine.name, year, quarter, num.death, num.dis, severe)
glm.result$prob <- glm.prediction
glm.result <- glm.result %>% mutate(
  pred = ifelse(
    prob > 0.5, TRUE, FALSE
  )
)
head(glm.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10)
head(glm.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 10)

glm.performance <- confusionMatrix(data = glm.result$pred, reference = glm.result$severe, positive = "TRUE")
print(glm.performance)	
```

### Conditional Logistic Regression

```{r fixed effect, include=FALSE}
# fixed effects on training and testing data ----
# train and test
set.seed(1)
data.len <- nrow(labeled.data.death)
train.indices <- sample(seq_len(nrow(labeled.data.death)), size = floor(data.len / 2))

labeled.data.death.train <- labeled.data.death[train.indices, ]
labeled.data.death.result <- labeled.data.death[-train.indices, ]
labeled.data.death.result <- labeled.data.death.result %>% filter(is.element(mine_id, unique(labeled.data.death.train$mine_id)))

fe.div.model <- clogit(severe ~ last.quarter.lost+last.year.lost+last.three.years.lost
                   +last.quarter.restrict+last.year.restrict+last.three.years.restrict
                   +last.quarter.viol+last.year.viol+last.three.years.viol
                   +last.quarter.death+last.year.death+last.three.years.death
                   + strata(mine_id), data = labeled.data.death.train)
summary(fe.div.model)

fe.div.prediction <- predict(object = fe.div.model, newdata = labeled.data.death.result,type = "lp") %>% sigmoid()
fe.div.result <- labeled.data.death.result %>% select(mine_id, mine.name, year, quarter, num.death, num.dis, severe)
fe.div.result$prob <- fe.div.prediction
fe.div.result <- fe.div.result %>% mutate(
  pred = ifelse(
    prob > 0.5, TRUE, FALSE
  )
)
head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10)
head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 10)

# confusion matrix and other performance metrics
#fe.div.performance <- confusionMatrix(data = fe.div.result$pred, reference = fe.div.result$severe, positive = "TRUE")
fe.div.performance <- table(fe.div.result$pred, fe.div.result$severe)
print(fe.div.performance)

# fixed effects on training and testing data (using labeled.data.lost) ----
# fe.div + num.days.lost: train and test
labeled.data.lost.train <- labeled.data.lost[train.indices, ]
labeled.data.lost.result <- labeled.data.lost[-train.indices, ]
labeled.data.lost.result <- labeled.data.lost.result %>% filter(is.element(mine_id, unique(labeled.data.lost.train$mine_id)))

fe.div2.model <- clogit(severe ~ last.quarter.lost+last.year.lost+last.three.years.lost
                       +last.quarter.restrict+last.year.restrict+last.three.years.restrict
                       +last.quarter.viol+last.year.viol+last.three.years.viol
                       +last.quarter.death+last.year.death+last.three.years.death
                       + strata(mine_id), data = labeled.data.lost.train)
summary(fe.div2.model)

fe.div2.prediction <- predict(object = fe.div2.model, newdata = labeled.data.lost.result,type = "lp") %>% sigmoid()
fe.div2.result <- labeled.data.lost.result %>% select(mine_id, mine.name, year, quarter, num.death, num.dis, severe)
fe.div2.result$prob <- fe.div2.prediction
fe.div2.result <- fe.div2.result %>% mutate(
  pred = ifelse(
    prob > 0.5, TRUE, FALSE
  )
)
# head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10)
# head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 10)

# confusion matrix and other performance metrics
fe.div2.performance <- confusionMatrix(data = fe.div2.result$pred, reference = fe.div2.result$severe, positive = "TRUE")
print(fe.div2.performance$table)
print(c(fe.div2.performance$overall["Accuracy"], fe.div2.performance$byClass[c("Sensitivity", "Specificity", "Precision", "F1")]))   
```

### Clustering

First, we cluster the data using K-means.

```{r clustering, include=FALSE}
# clustering ----
temp <- complete.active.quarters %>% filter(active==TRUE)
set.seed(20)
clustering <- kmeans(temp[, 6:ncol(temp)], 3, nstart=20)
freq_table <- table(clustering$cluster)

print(clustering$centers)
print(clustering$size)
```

Then, let's visualize the clusters.

```{r cluster visualization, include=FALSE}
clustering$cluster <- as.factor(clustering$cluster)
ggplot(temp, aes(num.death, viol.quantity, color = clustering$cluster)) + geom_point()
```

### Markov Chain

```{r markov, include=FALSE}
# assign low, mid, high risk labels
risk.labels <- c("low", "mid", "high")
cluster.to.risk.label <- sort.int(clustering$centers[,1], index.return = TRUE)$ix

# Melted format containing quarter-year, mine_id allowing for markov chain analysis ----
temp$cluster <- clustering$cluster
temp$date <- paste(as.character(temp$year), "-" , as.character(temp$quarter))
temp2 <- temp %>% select(mine_id, cluster, date)
melted <- melt(temp2, id=c("mine_id", "date"))
aggregate_cluster_seq <- cast(melted, mine_id~date)

# Markov Chain function ----
markov <- function(x){
  x <- as.integer(x)
  cluster_seq_raw <- x[!is.na(x)]
  if(length(cluster_seq_raw) > 1){
    cluster_sequence <- cluster_seq_raw[2:length(cluster_seq_raw)]
    cur_markov_chain <- createSequenceMatrix(cluster_sequence, sanitize = FALSE)
    return (list("mine_id" = x[1], 
                 "markov_chain" = cur_markov_chain, 
                 "num_rows" = length(cluster_seq_raw),
                 "low" = cluster.to.risk.label[1] %in% cluster_sequence, 
                 "mid" = cluster.to.risk.label[2] %in% cluster_sequence, 
                 "high" = cluster.to.risk.label[3] %in% cluster_sequence
    )
    )
  }
  return (list("mine_id" = x[1], 
               "markov_chain" = NA, 
               "num_rows" = 0,
               "low" = FALSE, "mid" = FALSE, "high" = FALSE
  )
  )
}

# create mine-date transition matrices ----
result<-apply(aggregate_cluster_seq, 1, markov)

# Create Overall transition matrix ----
markov_chains_ind <- sapply(result, "[[", 2)

#adding matrix of different dimentions, source: http://stackoverflow.com/questions/13571359/join-and-sum-not-compatible-matrices
add_matrices_1 <- function(...) {
  a <- list(...)
  cols <- sort(unique(unlist(lapply(a, colnames))))
  rows <- sort(unique(unlist(lapply(a, rownames))))
  out <- array(0, dim=c(length(rows), length(cols)), dimnames=list(rows,cols))
  for(M in a) { out[rownames(M), colnames(M)] <- out[rownames(M), colnames(M)] + M }
  return(out)
}

final_transition<- matrix(rep(0,times= 9), ncol = 3, nrow = 3)
colnames(final_transition) <- c(1,2,3)
rownames(final_transition) <- c(1,2,3)

for(i in 1:length(markov_chains_ind)){
  final_transition <- add_matrices_1(final_transition,markov_chains_ind[[i]])
}
print("Tranition matrix frequency")
final_transition

final_markov<- matrix(rep(0,times= 9), ncol = 3, nrow = 3)
for(i in 1:3){
  final_markov[i,] <- final_transition[i,]/sum(final_transition[i,])  
}

print("Transition matrix probability")
sorted.final.markov <- final_markov[cluster.to.risk.label, cluster.to.risk.label]
colnames(sorted.final.markov) <- risk.labels
rownames(sorted.final.markov) <- risk.labels
print(sorted.final.markov)
```

### SVM

We try to classify risky mines using SVM.

```{r svm, include=FALSE}
# check train and test data
head(labeled.data.lost.train)
head(labeled.data.lost.result)

# convert label from `bool` to `int`
labeled.data.lost.train <- labeled.data.lost.train %>% mutate(severe = severe*1)
labeled.data.lost.result <- labeled.data.lost.result %>% mutate(severe = severe*1)

svm.model <- svm(severe ~ 
                   last.quarter.lost+last.year.lost+last.three.years.lost
                 +last.quarter.restrict+last.year.restrict+last.three.years.restrict
                 +last.quarter.viol+last.year.viol+last.three.years.viol
                 +last.quarter.death+last.year.death+last.three.years.death,
                 data=labeled.data.lost.train, method="C-classification", kernel="linear")
summary(svm.model)

pred.train <- predit(svm.model, labeled.data.lost.train)
mean(pred.train==labeled.data.lost.train$severe)
```

## Evaluation